Cybersecurity impacts at the beginning\\ of Covid-19 pandemic in Italy

### 2021-03-04 article.tex
Drafting an introduction and other minor changes

\documentclass{easychair}
%\documentclass[EPiC]{easychair}
%\documentclass[EPiCempty]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

@@

% use this if you have a long article and want to create an index
% \usepackage{makeidx}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% Some of our commands for this guide.
%
\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

%\makeindex

%% Front Matter
%%
% Regular title as in the article class.
%
\title{The impacts of the beginning of the Covid-19 pandemic on the cybersecurity in Italy}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes
% into the list defined using \institute
%
\author{
Marco R. A. Bozzetti\inst{1}
\and
    Fausto Spoto\inst{2}
\and
   Luca Olivieri \inst{2}
}

@@

%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

@@

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair
\titlerunning{The Covid-19 pandemic impacts on the cybersecurity in Italy}

@@

OAD, the Digital Attacks Observatory in Italy, with the last 2020 edition \cite{OAD_report_2020} \footnote{The report is written in Italian, only 
the Executive Summary is in English: 186 A4 pages, 148 images and graphics, 11 Chapters (147 A4 pages) and 9 Attachments (39 A4 pages)} reaches the 
twelfth year of consecutive surveys on cyberattacks in Italy, and, as in the previous years, with the precious collaboration of the Italian Postal 
and Telecommunications Police. OAD is the only independent online survey in Italy on intentional digital attacks on IT systems of companies and public
bodies operating in Italy. The OAD survey does not provide for a predefined set of respondents, but it allows potential interested parties full and 
free access to the online questionnaire, in a totally anonymous manner. Thanks to the number of responses collected and their balanced distribution 
between companies and public bodies of various sizes and belonging to various product sectors, the OAD survey provides a specific picture of the cyberattacks in Italy.

The OAD 2020 survey took place during the Covid-19 pandemic, and it covered the entire 2019 and the first quarter of 2020, when in Italy the pandemic exploded.
The comparison of the data of 2019 with these of the first quarter 2020, provided by the same pool of respondents, has a statistical relevance and provides a 
clear indication of the Covid pandemic on the intentional cyber-attacks in Italy. This pandemic has been the trigger for a wide range of cyberattacks, mainly 
caused by the sudden - and in part unprepared - passage of many to agile working from home and to a strong use of every type of IT service on Internet, mainly
derived by the mobility lockdowns imposed by the Italian Authorities. The emerged respondents’ pool covers all the product sectors, including Public Administrations,
even if the majority of the respondents’ companies belong to the ICT sector (30.8\%). The 2020 pool  is well balanced for the size of the organizations, in terms
of number of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\% \footnote{According to the last ISTAT census}
of enterprises are SMEs, Small Medium Enterprises, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 
very small organizations, which are thee vast majority in Italy, and which are not normally considered in other cybersecurity surveys: 57.5\% of the OAD 2020 respondents
belong to structures with less than 250 employees, and of these 22.1\% under 10. The OAD questionaries refer for about half of the question to the cyber-attacks 
typologies and they main characteristics, if they have been relived. The other half of the questions refers to the technical and organizational security measures implemented
on the IT systems of the respondents. OAD therefore allows to contextualize the cyber-attacks with the cyber measures in production in each IT system considered. 
And in fact, in order to improve the motivation to complete the questionaries to all the potential anonymous respondents, OAD 2020 provided a macro qualitative evaluation
of the implemented measures, within the specific context of the company. The results of  this macro evaluation is shown in fig. 1, which shows that more than 2/3 of the respondents has a good level of cybersecurity.

%TODO: include fig 1

Fig. 2 compares the trend of the relived attacks from the beginning of the OAD survey in 2008: this comparison is only valid as a trend and not from a statistical point 
of view, due to the fact that the respondents pools year by year are different. Since 2007 to 2015, the “average trend” is focused about 40\% of the respondents that relived
cyber-attacks. (first red line in the figure). Since 2015 up to day, this average trend increased to 45\% of the respondents (second red line in the figure). And in 2018 OAD 
reached the pick of relived attacks, and  for the first time  the \% of relived attacks is higher that the \% of the not relived.

%TODO: include fig 2

@@

\begin{itemize}
\item As shown in fig. 1, the majority of the respondents , in the OAD 2020 survey, has a good level of cyber security: about 30\% are of the ICT sector. 
And therefore the IT Systems can properly react to the cyber attacks.

\item As pointed out before, 99,91\% of the companies in Italy are SME, and 95\% are under 10 employees. More r less is the same for the Public Administration (PA), 
the local ones in particular. And the number of respondents in the 2020 OAD survey is 57,7\% for SME/small P, and of these 22,1\% are under 10 employees. 
The hackers/crackers do not waist time for targeted attacks to very small realities.
\end{itemize}

@@

%TODO: include fig 3

@@

%TODO: include fig 4

@@

%TODO: include fig 5
%TODO: include fig 6
%TODO: include fig 7
%TODO: include fig 8

@@

To conclude, the OAD 2020 survey highlights a situation of cyberattacks of various types but all technically of ian high complexity and sophistication, with a slightly increasing 
spread  in Italy  compared to the trend that emerged in the twelve years of OAD surveys, but lower than the peak of 2018. Despite the proliferation of cyberattacks to varying 
degrees related to the Covid-19 pandemic, in the first four months of 2020 the general spread of attacks is at similar values to those of 2019: to be verified with the next OAD 2021
if there will be changes to this trend.

@@

>>>>

\documentclass{easychair}

@@

\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

\title{Cybersecurity impacts at the beginning\\ of Covid-19 pandemic in Italy}

\author{
Marco R. A. Bozzetti\inst{1}
\and
    Fausto Spoto\inst{2}
\and
   Luca Olivieri \inst{2}
}

@@

@@

\titlerunning{Cybersecurity impacts at the beginning of Covid-19 pandemic in Italy}

@@

The advent of the Covid-19 pandemic and the consequent lock-downs have led companies to face needs such as smart working, remotely work and digital business,
accelerating their digitalization efforts. As reported in \cite{Goasduff} by Gartner, a world's leading research and advisory company, before the pandemic crisis
most organizations moved their digital strategies forward at a steady pace, but during the Covid-19 pandemic there has been a significant increase in the 
development of digital products and services to maintain and accelerate customer engagement. However, the digitalization race has led to many cyber security issues
and the intensification of malicious cyber attacks around the world. In this scenario, Italy is an insteresting case of study because it is composed mostly from 
Small Medium Enterprises (SMEs) and therefore can highlight specific aspects instead of the various global views in which large companies have a greater influence.

\section{Digital Attacks Observatory}

The Digital Attacks Observatory (OAD)~\cite{OAD} is the only independent online survey in Italy about intentional digital attacks on IT systems of companies and public
bodies operating and in addition drawn up with the precious collaboration of the Italian Postal and Telecommunications Police. The OAD survey does not provide 
for a predefined set of respondents, but it allows potential interested parties full and free access to an online questionnaire, in a totally anonymous manner. 
Thanks to the number of responses collected and their balanced distribution between companies and public bodies of various sizes and belonging to various product 
sectors, the OAD survey provides a specific picture of the cyber-attacks in Italy.

\section{Data comparison}

OAD 2020 survey\footnote{The report is written in Italian, only the Executive Summary is in English: 186 A4 pages, 148 images and graphics, 11 Chapters
(147 A4 pages) and 9 Attachments (39 A4 pages)}~\cite{OAD_report_2020} took place during the Covid-19 pandemic, and it covered the entire 2019 and the first quarter 
of 2020, when in Italy the pandemic exploded. The data comparison of 2019 with these of the first quarter 2020, provided by the same pool of respondents, has a statistical relevance and provides a 
clear indication of the Covid pandemic on the intentional cyber-attacks in Italy. This pandemic has been the trigger for a wide range of cyberattacks, mainly 
caused by the sudden - and in part unprepared - passage of many to agile working from home and to a strong use of every type of IT service on Internet, mainly
derived by the mobility lockdowns imposed by the Italian Authorities. The emerged respondents’ pool covers all the product sectors, including Public Administrations,
even if the majority of the respondents’ companies belong to the ICT sector (30.8\%). The 2020 pool  is well balanced for the size of the organizations, in terms
of number of employees, between those below 250 and the largest ones. It must be taken into account that in Italy 99.91\%
\footnote{According to the last ISTAT census} %TODO: manca fonte/citazione
of enterprises are SMEs, under 250 employees, and 95\% are under ten employees. The OAD 2020 survey is therefore able to consider small and 
very small organizations, which are thee vast majority in Italy, and which are not normally considered in other cybersecurity surveys: 57.5\% of the OAD 2020 respondents
belong to structures with less than 250 employees, and of these 22.1\% under 10. The OAD questionaries refer for about half of the question to the cyber-attacks 
typologies and they main characteristics, if they have been relived. The other half of the questions refers to the technical and organizational security measures implemented
on the IT systems of the respondents. OAD therefore allows to contextualize the cyber-attacks with the cyber measures in production in each IT system considered. 
In fact, OAD 2020 provided a macro qualitative evaluation of the implemented measures (within the specific context of the company), in order to improve the motivation to 
complete the questionaries to all the potential anonymous respondents. The results of this macro evaluation is shown in fig. 1, which shows that more than $ 2/3 $ of the 
respondents has a good level of cybersecurity.

\begin{figure}
	\centering
		\includegraphics[width=1\textwidth]{pictures/fig1.png}
		\caption{}
		\label{fig:1}
\end{figure}

Fig. 2 compares the trend of the relived attacks from the beginning of the OAD survey in 2008: this comparison is only valid as a trend and not from a statistical point 
of view, due to the fact that the respondents pools year by year are different. Since 2007 to 2015, the “average trend” is focused about 40\% of the respondents that relived
cyber-attacks. (first red line in the figure). Since 2015 up to day, this average trend increased to 45\% of the respondents (second red line in the figure). In 2018 OAD 
reached the pick of relived attacks, and  for the first time  the \% of relived attacks is higher that the \% of the not relived.

\begin{figure}
	\centering
		\includegraphics[width=1\textwidth]{pictures/fig2.png}
		\caption{}
		\label{fig:2}
\end{figure}

@@

\begin{itemize}
\item As shown in fig. 1, the majority of the respondents , in the OAD 2020 survey, has a good level of cyber security: about 30\% are of the ICT sector. 
Therefore the IT Systems can properly react to the cyber attacks.

\item As pointed out before, 99,91\% of the companies in Italy are SME, and 95\% are under 10 employees. More r less is the same for the Public Administration (PA), 
the local ones in particular. The number of respondents in the 2020 OAD survey is 57,7\% for SME/small P, and of these 22,1\% are under 10 employees. 
The malicious hackers/crackers do not waist time for targeted attacks to very small realities.
\end{itemize}

@@

\begin{figure}
	\centering
		\includegraphics[width=1\textwidth]{pictures/fig3.png}
		\caption{}
		\label{fig:3}
\end{figure}

@@

\begin{figure}
	\centering
		\includegraphics[width=1\textwidth]{pictures/fig4.png}
		\caption{}
		\label{fig:4}
\end{figure}

@@

\begin{figure}
	\centering
		\includegraphics[width=1\textwidth]{pictures/fig5.png}
		\caption{}
		\label{fig:5}
\end{figure}
\begin{figure}
	\centering
		\includegraphics[width=1\textwidth]{pictures/fig6.png}
		\caption{}
		\label{fig:6}
\end{figure}
\begin{figure}
	\centering
		\includegraphics[width=1\textwidth]{pictures/fig7.png}
		\caption{}
		\label{fig:7}
\end{figure}
\begin{figure}
	\centering
		\includegraphics[width=1\textwidth]{pictures/fig8.png}
		\caption{}
		\label{fig:8}
\end{figure}

@@

The OAD 2020 survey highlights a situation of cyberattacks of various types but all technically of ian high complexity and sophistication, with a slightly increasing 
spread  in Italy  compared to the trend that emerged in the twelve years of OAD surveys, but lower than the peak of 2018. Despite the proliferation of cyberattacks to varying 
degrees related to the Covid-19 pandemic, in the first four months of 2020 the general spread of attacks is at similar values to those of 2019: to be verified with the next OAD 2021
if there will be changes to this trend.

@@

\bibliographystyle{plain}
\bibliography{biblio}
